Robust Bayesian Analysis of Causal Inference Problems

### 2024-03-21 JASA-template.tex
Notation for defs.
Some comments.

Now, to construct the joint likelihood function, we define an extended
output $2n\times 1$ column vector
$W\coloneqq\left(\begin{smallmatrix}Y \\ T^*\end{smallmatrix}\right)$
and corresponding $2n\times(2p+1)$ dimensional design matrix
\begin{align}
	Z &\coloneqq
        \begin{bmatrix}
           T_1 & X_1 & 0 \\
           \vdots & \vdots & 0 \\
           T_n & X_n & 0 \\
           0 & 0 & X_1 \\
           \vdots & \vdots & \vdots \\
           0 & 0 & X_n
        \end{bmatrix}
        =
	\begin{bmatrix}
		X_O & 0 \\
		0 & X_T
	\end{bmatrix}
\end{align}
where, $X_O = [T, X]$ and $X_T = X$. Then, considering the assumption of
Gaussian error terms, we have the following likelihood distribution
\begin{align}
	W\mid Z, \beta_T, \beta, \gamma, \sigma^2 \sim\normal\left(Z\nu, \Sigma\right)\label{eq:like:group},
\end{align}
where $\nu = (\beta_T, \beta^T, \gamma^T)^T$ and
\begin{align}
	\Sigma &=
	\begin{bmatrix}
		\sigma^2{I}_n & 0 \\
		0 & {I}_n
	\end{bmatrix}.
\end{align}

@@

Let $\pi_j$ denote the prior probability that the $j$-th
predictor is associated to the outcome or the 
treatment. That is, 
\begin{equation}
	\pi_j = P\left((\beta_j,\gamma_j)\not=(0,0)\right).
\end{equation}
Then we can define the following hierarchical model for spike and 
slab group LASSO so that,
for $1\le j\le p$,
\begin{align}
	(\beta_j,\gamma_j)^T \mid \pi_{j}, \sigma^2 &\sim 
	\pi_{j}\normal\left( \begin{bmatrix}
		0 \\
		0
	\end{bmatrix}, 
	\tau_1^2\begin{bmatrix}
		\sigma^2 & 0 \\
		0 & 1
	\end{bmatrix}\right)
	+ (1-\pi_{j}) \normal\left(\begin{bmatrix}
		0 \\
		0
	\end{bmatrix}, 
	\tau_0^2\begin{bmatrix}
		\sigma^2 & 0 \\
		0 & 1
	\end{bmatrix}\right)\\
	\beta_T\mid \sigma^2 &\sim \normal\left(0, \sigma^2\right)\\
	\frac{1}{\sigma^2}&\sim \text{Gamma}(a, b)\\
	\pi_{j} &\sim\text{Beta}\left(sq, s(1-q)\right).
\end{align}
In the hierarchical model, we fix sufficiently small $\tau_0$
$(1\gg\tau_0>0)$ so that  $(\beta_j, \gamma_j)$ has its probability mass 
concentrated around zero. Therefore, this represents the spike component of our prior specification. 
For the slab component, we consider $\tau_1$ to be large so that $\tau_1\ge 1$. This allows the prior for $(\beta_j,\gamma_j)$ to be flat, besides the spike component at the origin.. 
We illustrate the components of a bivariate spike and slab prior in 
\cref{fig:ssbl} (with fixed $\sigma=1$). We generate the spike component 
with $\tau_0=0.001$ and the slab component with $\tau_1=5$.

@@

For the precision term $1/\sigma^2$, a natural choice of prior is the gamma distribution
as it allows the control of both the location and the scale of the precision. To
ensure that the prior is able to represent the data, we consider $b=1$ and 
fix $a$ so that it represents the prior mean of the precision. In cases where we have no prior information,
we can simply consider a large value for $a$ so that the interval
$[0, 2a]$ contains the true value of the precision.
As defined earlier, $\pi_j$ is used as the selection probability
of the $j$-th predictor in either of the models and we use a beta prior to specify 
these selection probabilities where  $q_j$ represents our prior expectation of the selection probability ($\pi_j$) and $s$ acts as 
a concentration parameter.
For the causal effect, we want to use a Gaussian distribution that 
matches the scale of the noise term. Therefore, we consider $\beta_T\sim \normal(0,\sigma^2)$.

>>>>

Now, to construct the joint likelihood function, we define an extended
output $2n\times 1$ column vector
$W\coloneqq\left(\begin{smallmatrix}Y \\ T^*\end{smallmatrix}\right)$
and corresponding $2n\times(2p+1)$ dimensional design matrix
\begin{align}
	Z &\coloneqq
        \begin{bmatrix}
           T_1 & X_1 & 0 \\
           \vdots & \vdots & 0 \\
           T_n & X_n & 0 \\
           0 & 0 & X_1 \\
           \vdots & \vdots & \vdots \\
           0 & 0 & X_n
        \end{bmatrix}
        =
	\begin{bmatrix}
		X_O & 0 \\
		0 & X_T
	\end{bmatrix}
\end{align}
where, $X_O \coloneqq [T, X]$ and $X_T \coloneqq X$. Then, considering the assumption of
Gaussian error terms, we have the following likelihood distribution
\begin{align}
	W\mid Z, \beta_T, \beta, \gamma, \sigma^2 \sim\normal\left(Z\nu, \Sigma\right)\label{eq:like:group},
\end{align}
where $\nu \coloneqq (\beta_T, \beta^T, \gamma^T)^T$ and
\begin{align}
	\Sigma &\coloneqq
	\begin{bmatrix}
		\sigma^2{I}_n & 0 \\
		0 & {I}_n
	\end{bmatrix}.
\end{align}

@@

Let $\pi_j$ denote the prior probability that the $j$-th
predictor is associated with the outcome or the 
treatment. That is, 
\begin{equation}
	\pi_j \coloneqq P\left((\beta_j,\gamma_j)\not=(0,0)\right).
\end{equation}
Then we can define the following hierarchical model for spike and 
slab group LASSO so that,
for $1\le j\le p$,
\begin{align}
	(\beta_j,\gamma_j)^T \mid \pi_{j}, \sigma^2 &\sim 
	\pi_{j}\normal\left( \begin{bmatrix}
		0 \\
		0
	\end{bmatrix}, 
	\tau_1^2\begin{bmatrix}
		\sigma^2 & 0 \\
		0 & 1
	\end{bmatrix}\right)
	+ (1-\pi_{j}) \normal\left(\begin{bmatrix}
		0 \\
		0
	\end{bmatrix}, 
	\tau_0^2\begin{bmatrix}
		\sigma^2 & 0 \\
		0 & 1
	\end{bmatrix}\right)\\
	\beta_T\mid \sigma^2 &\sim \normal\left(0, \sigma^2\right)\\
	\frac{1}{\sigma^2}&\sim \text{Gamma}(a, b)\\
	\pi_{j} &\sim\text{Beta}\left(sq, s(1-q)\right).
\end{align}
In the hierarchical model, we fix sufficiently small $\tau_0$
$(1\gg\tau_0>0)$ so that  $(\beta_j, \gamma_j)$ has its probability mass 
concentrated around zero. Therefore, this represents the spike component of our prior specification. 
For the slab component, we consider $\tau_1$ to be large so that $\tau_1\ge 1$. This allows the prior for $(\beta_j,\gamma_j)$ to be flat beyond the spike component at the origin. 
We illustrate the components of a bivariate spike and slab prior in 
\cref{fig:ssbl} (with fixed $\sigma=1$). We generate the spike component 
with $\tau_0=0.001$ and the slab component with $\tau_1=5$.

@@

For the precision term $1/\sigma^2$, a natural choice of prior is the gamma distribution
as it allows the control of both the location and the scale of the precision.
%%% MT to TB: why not simply say the prior is exponential?
To ensure that the prior is able to represent the data, we consider $b=1$ and 
fix $a$ so that it represents the prior mean of the precision.
%%% MT to TB: "no prior information" seems a bit misleading here? I suggest removing
% In cases where we have no prior information,
We can simply consider a large enough value for $a$
%%% MT to TB: the next phrase is confusing... do you mean to say "with probability ..." (not sure what that is for [0,2a]))
so that the interval
$[0, 2a]$ contains the true value of the precision.
As defined earlier, $\pi_j$ is used as the selection probability
of the $j$-th predictor in either of the models and we use a beta prior to specify 
these selection probabilities where  $q_j$ represents our prior expectation of the selection probability ($\pi_j$) and $s$ acts as 
a concentration parameter.
For the causal effect, we want to use a Gaussian distribution that 
matches the scale of the noise term. Therefore, we consider $\beta_T\sim \normal(0,\sigma^2)$.
